### 2012-02-22 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

We want to capture the following types.

@@

\todo[inline]{Doplnit PathType o boolean for bound, volania zoznam funkcii volanych na tuto path.}

@@

\todo[inline]{ALG Extraction of a Type from TypeNode nie uplne sedi s definiciou typov. DP JS ma nieco navyse.}

@@

\todo[inline]{ALG \ref{ALG_func_determineExpressionType}: Vymysliet a dopisat ako ziskavat typ z PathExprNode. Bude potrebne vyskusat implementaciu, ako sa chova.}

@@

\STATE TODO
\end{algorithmic}
\end{algorithm}

@@

\STATE TODO
\end{algorithmic}
\end{algorithm}

@@

\STATE TODO
\end{algorithmic}
\end{algorithm}

@@

    \FORALL{$bindingNode \in node.getChild(TupleStreamNode).getChildren()$}
        \IF{$is(bindingNode, ForClauseNode)$}
            \STATE $expr = bindingNode.getChild(BindingSequenceNode).$ $getChild(ExprNode)$
            \IF{$getType(expr) == PathExpr$ without predicates except the last step, using only child and descendant axes}
                \STATE add $bindingNode.varName$ to $forVars$
            \ENDIF
        \ENDIF
    \ENDFOR

>>>>

But firstly, we describe types of expressions we want to capture and their features.

\subsection{Expression Types}

@@

\subsection{PathType}
$PathType$ contains additional information. The represented path is contained by a list of its steps, in particular instances of $StepExprNode$. If a step is a reference to a variable which type is $PathType$, we also want to include this information. Therefore, $PathType$ contain association between the steps and other $PathTypes$ and this association is defined for the $PathType$ variable steps.

To differ between a common $PathType$ selecting a set of nodes and a $PathType$ bound to a for variable in a FLWOR expression, $PathType$ structure contains a flag $isForBound$.

Additionaly, $PathType$ contains a list of special functions that were called with an argument of $PathType$ type. The motivation is that in some cases of function calls, we want to know that the function call is performed with an instance of $PathType$ because then, we can determine a type of the function call more precisely. Those special functions are built-in functions $data$, $min$, $max$ \todo{Doplnit}.

In summary, we represent $PathType$ as a structure with the following member variables. \todo{Radsej UML?}
\begin{itemize}
\item $stepNodes$ - A list of references to $StepExprNode$ instances.
\item $forBoundSubsteps$ - An association between references from $stepNodes$ and instances of $PathType$ type.
\item $isForBound$ - Boolean flag determining if the type was bound to a for variable in a for clause.
\item $specialFunctionCalls$ - List of special functions called with this instance as an argument.
\end{itemize}

\subsection{Cardinality of Types}

@@

\todo[inline]{ALG Extraction of a Type from TypeNode nie uplne sedi s definiciou typov. DP JS ma nieco navyse. A tento alg je nedokonceny!!}

@@

@@

\STATE $stepNodes =$ an empty array
\STATE $forBoundSubsteps =$ an empty associative array
\FORALL{$stepNode \in pathExprNode.getChildren()$}
	\STATE $detailNode = stepNode.getChild(DetailNode)$
	\IF{$detailNode != null$}
		\IF{$is(detailNode, VarRefNode)$}
			\STATE $type = getType(detailNode)$ // $PathType$
			\IF{$type.isForBound$}
				\STATE $forBoundSubsteps[stepNode] = type$
				\STATE add $stepNode$ to $stepNodes$
			\ELSE
				\STATE add all $type.stepNodes$ to $stepNodes$
			\ENDIF
		\ELSE
			\STATE add $stepNode$ to $stepNodes$
		\ENDIF
	\ELSE
		\STATE add $stepNode$ to $stepNodes$
	\ENDIF
\ENDFOR

\RETURN $PathType$ with $stepNodes$, $forBoundSubsteps$, and $isForBound$ set to $false$
\end{algorithmic}
\end{algorithm}

@@

\IF{$type$ is $UnknownType$}
	\RETURN $UnknownType$
\ELSIF{$type$ is $NodeType$ or XSD atomic type}
	\RETURN $type$ with cardinality set to exactly-one
\ELSE
	\STATE // $PathType$
	\RETURN $type$ with $isForBound$ flag set to $true$
\ENDIF
\end{algorithmic}
\end{algorithm}

@@

\IF{$type$ is $UnknownType$}
	\RETURN $UnknownType$
\ELSIF{$type$ is $NodeType$ or XSD atomic type}
	\RETURN $type$ with cardinality set to zero-or-more
\ELSE
	\STATE // $PathType$
	\RETURN $type$ with $isForBound$ flag set to $false$
\ENDIF
\end{algorithmic}
\end{algorithm}

@@

    \FORALL{$bindingNode \in node.getChild(TupleStreamNode).getChildren()$}
        \IF{$is(bindingNode, ForClauseNode)$}
            \STATE $expr = bindingNode.getChild(BindingSequenceNode).$ $getChild(ExprNode)$
            \IF{$getType(expr) == PathType$ using only child and descendant axes}
                \STATE add $bindingNode.varName$ to $forVars$
            \ENDIF
        \ENDIF
    \ENDFOR
